%%%%%%%%%%%%%%%%%%%%%%%% referenc.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% sample references
% %
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{99.}%
	
	
	
	\bibitem{cisco} Cisco, ``{Cisco visual networking index: Global mobile data traffic forecast up-date 2015-2019},'' Online, whitepaper, February 2016. [Online]. Available: \url{{http://www.cisco.com/c/en/us/solutions/collateral/service-provider/visual-networking-index-vni/mobile-white-paper-c11-520862.html}}
	
	\bibitem{lteu-forum}``{LTE-U technical report: Coexistence study for LTE-U SDLV1.0 (2015-02)},''Online, LTE-U Forum, Tech. Rep., February 2015. [Online]. Available: \url{http://www.lteuforum.org/uploads/3/5/6/8/3568127/lte-u\_forum\_lte-u\_technical\_report\_v1.0.pdf}
	
	\bibitem{3gpp}{3GPP}, ``Feasibility study on licensed-assisted access to unlicensed spectrum,'' \emph{{TR 36.889 v13.0.0}}, July 2015.
	
	
	\bibitem{qual}``{LTE in unlicensed spectrum: Harmonious coexistence with Wi-Fi},'' Online, Qualcomm, Inc, White Paper, June 2014. [Online]. Available: \url{{https://www.qualcomm.com/documents/lte-unlicensed-coexistence-whitepaper}}
	
	\bibitem{zhang} {R. Zhang, M. Wang, et al}, ``Lte-unlicensed: The future of spectrum aggregation for cellular networks,'' \emph{IEEE Wireless Communications}, vol.~22, no.~3, pp. 150--159, June 2015.
	
	\bibitem{timo} {T. Nihtil{\"a}, V. Tykhomyrov, et al}, ``System performance of lte and ieee 802.11 coexisting on a shared frequency band,'' in \emph{Wireless Communications and Networking Conference , 2013 IEEE}, Shanghai, April 2013,pp. 1038--1043.
	
	\bibitem{google} ``{Comments of {G}oogle, {I}nc. In the Matter of the Office of Engineering and Technology and Wireless Telecommunications Bureau seeks Information on Current Trends in LTE-U and LAA Technology},'' before the FCC, June 2015, docket no. 15-105.
	
	\bibitem{80211}``{IEEE Standard for information technology--Telecommunications and information exchange between systems Local and metropolitan area networks--Specific requirements Part 11: Wireless LAN Medium Access Control (MAC) and Physical Layer (PHY) Specifications},'' \emph{{IEEE Std 802.11-2012 (Revision of IEEE Std 802.11-2007)}}, pp. 818--972, March 2012.
	
	\bibitem{tsi} ``{LTE in a Nutshell},'' Online, Telesystem Innovations Inc., White Paper, 2010.
	
	\bibitem{dulaimi} {A. Al-Dulaimi, S. Al-Rubaye, et al}, ``5g communications race: Pursuit of more capacity triggers lte in unlicensed band,'' \emph{IEEE Vehicular Technology Magazine}, vol.~10, no.~1, pp. 43--51, March 2015.
	
	\bibitem{qual2} ``{Comments of {Q}ualcomm, {I}nc. In the Matter of the Office of Engineering and Technology and Wireless Telecommunications Bureau seeks Information on Current Trends in LTE-U and LAA Technology},'' before the FCC, June 2015, docket no. 15-105.
	
	\bibitem{abinader} {F. M. Abinader, E. P. L. Almeida, et al}, ``Enabling the coexistence of lte and wi-fi in unlicensed bands,'' \emph{IEEE Communications Magazine}, vol.~52, no.~11, pp. 54--61, November 2014.
	
	\bibitem{jian} {Y. Jian, {C.-F} Shih, et al}, ``Coexistence of wi-fi and laa-lte: Experimental evaluation, analysis and insights,'' in \emph{Communication Workshop, 2015 IEEE Int. Conf.}, London, June 2015, pp. 2325--2331.
	
	\bibitem{bhorkar} A.~Bhorkar, C.~Ibars, and P.~Zong, ``On the throughput analysis of lte and wifi in unlicensed bands,'' in \emph{Signals, Systems and Computers, 2014 48th Asilomar Conf.}, Pacific Grove, CA, November 2015, pp. 1309--1313.
	
	\bibitem{cablelabs}J.~Padden, ``{Wi-Fi vs. duty cycled LTE: A balancing act},'' Online, CableLabs, White Paper. [Online]. Available: \url{{http://www.cablelabs.com/wi-fi-vs-duty-cycled-lte/}}
	
	\bibitem{cano2} C.~Cano and D.~J. Leith, ``Coexistence of wifi and lte in unlicensed bands: A proportional fair allocation scheme,'' in \emph{Communication Workshop, 2015 IEEE Int. Conf.}, London, June 2015, pp. 2288--2293.
	
	\bibitem{li} Y.~Li, J.~Zheng, and Q.~Li, ``Enhanced listen-before-talk scheme for frequency reuse of licensed-assisted access using lte,'' in \emph{Personal, Indoor, and Mobile Radio Communications, 2015 IEEE 26th Annu. Int. Symp.}, Hong Kong, August 2015, pp. 1918--1923.
	
	\bibitem{kwan} {R. Kwan, R. Pazhyannur, et al}, ``Fair co-existence of licensed assisted access lte (laa-lte) and wi-fi in unlicensed spectrum,'' in \emph{Computer Science and Electronic Engineering Conf., 2015 7th}, Colchester, United Kingdom, September 2015, pp. 13--18.
	
	\bibitem{cano1} C.~Cano and D.~J. Leith, ``{Unlicensed LTE/WiFi coexistence: Is LBT inherently fairer than CSAT?}'' November 2015, {ArXiv pre-print, Computer Science - Networking and Internet Architecture, arXiv:1511.06244}.
	
	\bibitem{yoon} {J. Yoon, S. Yun, et al}, ``{Maximizing Differentiated Throughput in IEEE 802.11e Wireless LANs},'' in \emph{{ Proceedings of the 31st IEEE Conference on Local Computer Networks}}, vol.~14, no.~6, November 2006, pp. 411--417.
	
	\bibitem{chou} C.~T. Chou, K.~G. Shin, and S.~Shankar, ``{Contention-based Airtime Usage Control in Multirate IEEE 802.11 Wireless LANs},'' \emph{{IEEE/ACM Transactions on Networking}}, vol.~14, no.~6, pp. 1179--1192, December 2006.
	
	\bibitem{vu} H.~L. Vu and T.~Sakurai, ``{Collision probability in saturated IEEE 802.11 networks},'' in \emph{{Australian Telecommunication Networks \& Applications Conference}}, September 2006, pp. 1--5.
\end{thebibliography}
