%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
%\motto{Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
\chapter{U-LTE and Wi-Fi Coexistence: A Survey}
\label{survey} % Always give a unique label
% use \chaptermark{}
% to alter or adjust the chapter heading in the running head

\abstract*{Coexistence between \mbox{U-LTE} and \mbox{Wi-Fi} networks is a deciding factor on the acceptance of \mbox{U-LTE}. As a result, a large number of studies have been carried out to identify what could be the effects that \mbox{U-LTE} may cause to \mbox{Wi-Fi} and which mechanisms could be used to ensure that these two technologies share the $5$ GHz unlicensed frequency band in an efficient and fair manner. This chapter presents a survey on related work to answer the following questions: (i) what issues arise from simultaneous operation of LTE and \mbox{Wi-Fi} in the same spectrum bands, (ii) which technology is affected the most, and (iii) which factors determine the impacts of \mbox{U-LTE} to \mbox{Wi-Fi}. It also identifies the strengths and weaknesses of existing solutions and suggests potential strategies to improve performance of these two technologies.}

Coexistence between \mbox{U-LTE} and \mbox{Wi-Fi} networks is a deciding factor on the acceptance of \mbox{U-LTE}. As a result, a large number of studies have been carried out to identify what could be the effects that \mbox{U-LTE} may cause to \mbox{Wi-Fi} and which mechanisms could be used to ensure that these two technologies share the $5$ GHz unlicensed frequency band in an efficient and fair manner. This chapter presents a survey on related work to answer the following questions: (i) what issues arise from simultaneous operation of LTE and \mbox{Wi-Fi} in the same spectrum bands, (ii) which technology is affected the most, and (iii) which factors determine the impacts of \mbox{U-LTE} to \mbox{Wi-Fi}. It also identifies the strengths and weaknesses of existing solutions and suggests potential strategies to improve performance of these two technologies.

\section{Impacts of U-LTE on Wi-Fi Operation and Performance}

In \cite{original-LTE-Wi-Fi-VTC-2013}, extensive simulations were performed to assess the performance of LTE and \mbox{Wi-Fi} when the technologies are coexisting in an office environment. Both single-floor and multi-floor office environments, with different assumptions as to the density of \mbox{Wi-Fi} and LTE nodes, have been considered. The simulation results show that, in the absence of any modification to the LTE channel access mechanism, channel sharing between LTE and \mbox{Wi-Fi} networks is significantly unfair for \mbox{Wi-Fi} networks. While LTE experience only marginal performance loses when \mbox{Wi-Fi} is present on the same band (about $4$\% from the baseline performance), in a sparse deployment of $1$ AP per system per floor, \mbox{Wi-Fi} could lose up to $70$\% compared to its baseline performance.  In a dense deployment of $5$ APs per system per floor, the losses seen by \mbox{Wi-Fi} were nearly $100$\%. Detailed investigations in \cite{original-LTE-Wi-Fi-VTC-2013} indicate that for \mbox{Wi-Fi} the channel is blocked when LTE interference is present, and thus \mbox{Wi-Fi} nodes remain in ``listen'' for a clear channel mode most of the time.

The authors in \cite{original-LTE-Wi-Fi-WCNC-2013} present observations similar to those in \cite{original-LTE-Wi-Fi-VTC-2013} on the effects of unmodified LTE on the performance of \mbox{Wi-Fi} networks sharing the same frequency band. Specifically, it was found that as network load is increased LTE performance suffers only a minor degradation while \mbox{Wi-Fi} performance drops significantly. These results can be explained by the increasing LTE occupancy in the shared band because LTE does not follow the same rules as \mbox{Wi-Fi} in shared medium access. When there is ongoing transmission on the channel, \mbox{Wi-Fi} politely defers its transmissions while LTE will always choose to transmit and simply select a more robust transmission mode by adapting its modulation and channel coding scheme in order to cope with the higher levels of interference. This aggressive behavior quickly results in a situation where LTE terminals take all transmission opportunities while \mbox{Wi-Fi} devices are locked in defer and back-off procedures. Fortunately, the results in \cite{original-LTE-Wi-Fi-WCNC-2013} have also demonstrated that the severity of this negative impact on \mbox{Wi-Fi} can be efficiently controlled by restricting LTE activity.

The authors in \cite{LTE-U-PIMRC-2014} analyze the performance degradation of \mbox{Wi-Fi} in the presence of \mbox{LTE-U} with the probability of \mbox{Wi-Fi} accessing the channel is used as the main metric. Numerical results indicate that \mbox{Wi-Fi} is negatively affected by conventional LTE operation due to LTE's almost continuous transmissions subsequently blocking \mbox{Wi-Fi} access. Specifically, given the two modes of operations currently proposed for \mbox{LTE-U} in the unlicensed spectrum, the ``off'' period presented by the LTE protocol is too short for \mbox{Wi-Fi} users to access the channel. As a result, \mbox{Wi-Fi} is at risk of spending a significant amount of time in the ``listening'' mode when LTE transmissions are present in the same channel.

The work in \cite{U-LTE-Google-WP} presents initial investigations on the coexistence of two versions of license-anchored \mbox{U-LTE} (i.e., \mbox{LTE-U} and \mbox{LAA-LTE}) and \mbox{Wi-Fi} in $5$ GHz frequency band. Results show that \mbox{LTE-U} poorly coexists with \mbox{Wi-Fi} primarily due to two factors: (i) the incompatibility of \mbox{LTE-U}'s duty-cycling mechanism with \mbox{Wi-Fi} equipment and (ii) the lack of an effective coexistence mechanism in scenarios where \mbox{LTE-U} and \mbox{Wi-Fi} devices hear each other at moderate but non-negligible power levels. Additionally, \mbox{LAA-LTE} with LBT does not by itself guarantee successful fair coexistence with \mbox{Wi-Fi}. The results in \cite{U-LTE-Google-WP} were submitted to the FCC in June 2015 to demonstrate that, although any wireless technology should have the ability to utilize unlicensed spectrum within the FCC's rules, \mbox{U-LTE} has the potential to crowd out unlicensed services.

An experiment-based study on the effect of \mbox{LTE-U} to \mbox{Wi-Fi} is presented in \cite{LTE-U-CableLabs}. The LTE signal level was set higher than the \mbox{Wi-Fi} clients' LBT energy detection threshold (i.e., when LTE is on, the \mbox{Wi-Fi} client should sense their presence and not transmit). \mbox{Wi-Fi} throughput and latency were measured when data is transmitted through the \mbox{Wi-Fi} network with varying duty cycles and periods of LTE signals. The results in \cite{LTE-U-CableLabs} indicate that, as expected, increasing the \mbox{LTE-U} duty cycle degrades both \mbox{Wi-Fi} throughput and latency performance since it decreases \mbox{Wi-Fi} transmission opportunities accordingly. If the duty cycle period is too high, \mbox{Wi-Fi} latency is negatively impacted (while \mbox{Wi-Fi} throughput is nearly unchanged, given the same duty cycle) since \mbox{Wi-Fi} frames have to be buffered during long LTE ``on'' period. However, if the duty cycle period is configured as too low (e.g., $10$ msec), \mbox{Wi-Fi} throughput degrades due to the fact that LTE ``on'' and ``off'' periods are too short for \mbox{Wi-Fi} users to both access the channel and complete their transmissions. Furthermore, the authors in \cite{LTE-U-CableLabs} indicate that \mbox{LTE-U} duty cycle cannot strictly results in corresponding air time and throughput sharings. For example, with a duty cycle of $50$\%, \mbox{LTE-U} is likely to capture more than $50$\% of the channel resources. The reason is that when \mbox{LTE-U} starts its transmissions (regardless of ongoing \mbox{Wi-Fi} frame transmissions), many \mbox{Wi-Fi} frames are corrupted. The resulting transmission failures lead to multiple frame re-transmissions and, more importantly, mistakenly force \mbox{Wi-Fi} transceivers to operate at lower rates (in this case, lowering the channel coding and modulation modes is not necessary and is a waste of channel efficiency).


\section{Existing Solutions to Address U-LTE and Wi-Fi Coexistence Concerns}

Various coexistence mechanisms proposed for \mbox{U-LTE} are surveyed in \cite{U-LTE-survey-2014, U-LTE-survey-2015, U-LTE-5G-2015}. The proposed mechanisms include: Dynamic Channel Selection (DCS), transmission power control, opportunistic secondary cell ``off'', carrier-sense adaptive transmission (CSAT) (in \mbox{LTE-U}), and LBT (in \mbox{LAA-LTE}). When \mbox{U-LTE} and \mbox{Wi-Fi} share the common $5$ GHz radio frequency band, those mechanisms are found to be useful to reduce RFI and improve the spectrum utilization efficiency. However, the roles and performance of each varies greatly depending on network and system parameters including network scale, node density, deployment and radio environment (i.e., indoor, outdoor, short range, long range, etc.), and network load profiles. 

In order to see how LBT mechanisms employed by \mbox{LAA-LTE} can help promote fair coexistence, a simulation-based study is carried out and reported in \cite{LBT-CableLabs-2014}. LBE LBT specified by ETSI \cite{LBT-ETSI-2014} and IEEE 802.11e Enhanced Distributed Channel Access (EDCA) are assumed for \mbox{LAA-LTE} and \mbox{Wi-Fi}, respectively. The most important observation from \cite{LBT-CableLabs-2014} is that LBT compliant to ETSI regulation is not sufficient for fair coexistence: \mbox{Wi-Fi} STAs have much lower probability of successful channel access compared to \mbox{LAA-LTE} users. One major reason for this phenomenon is the non-exponential back-off employed by \mbox{LAA-LTE} LBT. Unfortunately, no form of exponential back-off LBT is studied in \cite{LBT-CableLabs-2014}.

In \cite{U-LTE-Wi-Fi-Qualcomm-2014, LTE-U-Wi-Fi-LTE-U-Forum-2015, LTE-U-Qualcomm-2015, U-LTE-Wi-Fi-Qualcomm-FCC-2015}, the performance of \mbox{LTE-U} and \mbox{LAA-LTE} and \mbox{Wi-Fi} in a shared frequency band is evaluated. DCS and opportunistic secondary cell ``off'' in unlicensed spectrum (\mbox{U-LTE} small cells would release the unlicensed carriers and fall back to the anchor carrier in licensed spectrum at low traffic load) are jointly used with CSAT and LBT. The results show that coexistence has a negative but controllable impact on \mbox{Wi-Fi} performance. In \cite{U-LTE-Wi-Fi-Qualcomm-2014, LTE-U-Wi-Fi-LTE-U-Forum-2015, LTE-U-Qualcomm-2015}, \mbox{LTE-U} can be a better neighbor to \mbox{Wi-Fi} than \mbox{Wi-Fi} would be to itself, in some scenarios. The underlying design that allows \mbox{LTE-U} to achieve high spectral efficiency while being a good neighbor to \mbox{Wi-Fi} is achieved through a set of carefully designed coexistence techniques, including DCS, secondary cell ``duty cycle'' in unlicensed spectrum (i.e., CSAT), and opportunistic secondary cell ``off'' in unlicensed spectrum. Specifically, in scenarios where the density of \mbox{Wi-Fi} APs and small cells is low or moderate, DCS and opportunistic secondary cell ``off'' are sufficient to meet the coexistence requirement. When \mbox{LTE-U} devices replace \mbox{Wi-Fi} devices, they can achieve significantly higher throughputs due to their high spectral efficiency. In addition, the performance of neighboring \mbox{Wi-Fi} is unchanged or even slightly improved since \mbox{LTE-U} devices can finish transmissions faster, and incur less interference, the similarly deployed \mbox{Wi-Fi} APs. However, if the density of \mbox{Wi-Fi} devices and \mbox{LTE-U} small cells is high, DCS and opportunistic secondary cell ``off'' alone cannot guarantee harmonious coexistence with \mbox{Wi-Fi} and therefore CSAT or LBT is required. Resutls in \cite{LTE-U-Wi-Fi-LTE-U-Forum-2015, U-LTE-Wi-Fi-Qualcomm-FCC-2015} were submitted to the FCC in 2015 to support \mbox{U-LTE} technologies.

A systematic and large-scale network-wide study of \mbox{LAA-LTE} and \mbox{Wi-Fi} performance in a wide range of realistic deployment scenarios and network densities in the unlicensed $5$ GHz band is presented in \cite{LTE-U-ICC-WS-2015}. The simulation results in all considered coexistence scenarios demonstrate that both \mbox{LAA-LTE} and \mbox{Wi-Fi} significantly benefit from the large number of available channels and the isolation provided by building shielding at $5$ GHz. They also suggest that deploying \mbox{LAA-LTE} with a random channel selection scheme is feasible for lower network densities. For typical indoor deployments of high density, implementing \mbox{LTE-U} interference-aware channel selection with respect to \mbox{Wi-Fi} is superior to LBT in terms of achieved throughput for both technologies. Additionally, LBT can increase \mbox{LAA-LTE} user throughput when multiple outdoor \mbox{LAA-LTE} networks deployed by different cellular operators coexist.

The work in \cite{Enhanced-LTE-U-thesis-2015} investigates the behavior and performance of two existing LBT mechanisms that are designed following the coexistence standard specified by ETSI, namely LBE- and FBE-based mechanisms \cite{LBT-ETSI-2014}. The Jain's fairness index has been used to assess the coexistence of \mbox{LAA-LTE} using these two LBT mechanisms and compare to \mbox{Wi-Fi} using CSMA-CA. The simulations in \cite{Enhanced-LTE-U-thesis-2015} show that the FBE-based mechanism, using a fixed contention window, impairs the channel access opportunity of \mbox{Wi-Fi}'s CSMA-CA using an adaptive contention window. The simulations reveal that FBE-based mechanism tends to aggressively occupy the channel. In some cases, \mbox{Wi-Fi} devices are starved with very few, or even no, chances to access the channel. This poor fairness is mainly caused by the short CCA sensing period of FBE-based mechanism. CCA is applied only once and then FBE-based mechanism may start its transmission immediately while LBE and \mbox{Wi-Fi}-based mechanisms are still decrementing their respective back-off counters. The fairness is worsened by the longer frames used by FBE. Another observation is that, again due to equal CCA sensing time, when multiple FBE-based equipment are contending for the channel, they are prone to serious collisions (if they are accidentally synchronized) or suffer a significant unfairness (if they are asynchronous). To cope with those issues, tuning the values of the back-off scaler ($q$) to extend the contention window size and using CCA procedure similar to that of LBE-based mechanism have been suggested for LBE and FBT-based mechanism, respectively. The results in \cite{Enhanced-LTE-U-thesis-2015} demonstrate that the modified LBE-based mechanism still cannot sufficiently improve the level of fairness achieved with others. This could be because simply empirically tuning the back-off scaler, while keeping the CCA principle unchanged, cannot compensate for the exponential growth of window size adopted by \mbox{Wi-Fi}'s CSMA-CA. The modified FBE-based mechanism can offer better fairness when coexisting with \mbox{Wi-Fi}.

A comparison of \mbox{LTE-U} and \mbox{LAA-LTE} is presented in \cite{LBT-CSAT-2015}. The analysis in \cite{LBT-CSAT-2015} shows that for sufficiently long LTE transmission times, the LTE throughputs achieved by CSAT and LBE are almost identical. However, for shorter LTE transmission times, \mbox{LTE-U} provides lower LTE throughput than \mbox{LAA-LTE} due to higher LTE/\mbox{Wi-Fi} collision probability of \mbox{LTE-U}. Besides, while shorter LTE transmission time decreases the tail of the \mbox{Wi-Fi} delay distribution, the percentage of packets that suffer from long delays increases. The results also indicate that when appropriately configured, \mbox{LTE-U} and \mbox{LAA-LTE} provide the same level of fairness to \mbox{Wi-Fi}. The selection of coexistence mechanisms is primarily driven by the operator's interests that include implementation complexity, LTE throughput, operational and management costs as well as strategic decisions on targeted markets.

Coordinated coexistence between \mbox{U-LTE} and \mbox{Wi-Fi} is investigated in \cite{U-LTE-5G-2015, Coordinated-LTE-U-Wi-Fi-2015}. The authors in \cite{U-LTE-5G-2015} propose a method of centralized system management which combines \mbox{LTE-U} and \mbox{Wi-Fi} through network function virtualization (NFV) interconnections. This technique may enable seamless transfers of resources between \mbox{LTE-U} and \mbox{Wi-Fi} using in-the-cloud control of distributed APs. However, only \textit{conceptual} network architectures and mechanisms are presented. The authors in \cite{Coordinated-LTE-U-Wi-Fi-2015} present a Software Defined Networking (SDN) architecture to support logically-centralized dynamic spectrum management involving multiple autonomous networks to improve spectrum utilization and facilitate coexistence. The basic design goal is to support the seamless communication and information dissemination required for coordination of heterogeneous networks. The system consists of two-tiered controllers mainly responsible for the control plane. The Global Controller (GC) acquires and processes global network state information (radio coverage maps, coordination algorithms, policy and network evaluation matrices, etc.) and controls the flow of information between Regional Controllers (RCs) and databases based on authentication and other regulatory policies. RCs acquire local visibility needed for radio resource allocation at wireless devices (device location, frequency band, duty cycle, power level, and data rate, etc.). Joint transmission power control and time division channel access optimizations are further proposed. Analytical results in \cite{Coordinated-LTE-U-Wi-Fi-2015} demonstrate that, \textit{with full buffer traffic assumption}, centralized optimization approaches can provide fair access to the spectrum for \mbox{LTE-U} and \mbox{Wi-Fi} networks.

An experimental evaluation of \mbox{U-LTE} interference effects on \mbox{Wi-Fi} performance under various network conditions, along with some suggestions for better coexistence of \mbox{U-LTE} and \mbox{Wi-Fi} networks, are presented in \cite{LTE-U-Experiment-ICC-WS-2015}. Various system parameters (bandwidth, center frequency, etc.) are swept to identify the most significant parameters that determine the levels of LTE interference introduced to \mbox{Wi-Fi} carrier sense and their effects on performance. The results indicate that \mbox{Wi-Fi} throughput can be heavily degraded by \mbox{LAA-LTE} transmissions with $3$/$5$/$10$ MHz bandwidth (especially $3$/$5$ MHz). Besides, \mbox{LAA-LTE} transmissions can have small impact on \mbox{Wi-Fi} throughput when using a $1.4$ MHz channel with center frequencies located on the guard bands or the center frequencies of \mbox{Wi-Fi} channels. However, the authors in \cite{LTE-U-Experiment-ICC-WS-2015} do not clearly define what \mbox{LAA-LTE} really mean in their work. It seems to be that they simply perform experiments with conventional LTE transceivers of varying power spectral densities and do not incorporate any coexistence mechanism into the LTE system.


\input{referenc-survey}